KidsAIR wood stove grades and usage

### 2020-12-24 stove_grade_usage_cleaning.Rmd
filter compliance data updates

kids_kw_report_clean <- kids_kw_report %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
  arrange(area, home_winter_id) %>% 
  mutate(filter_type = fct_collapse(filter_type,
                                    "Honeywell" = "Honeywell",
                                    "Large Filtrete" = "Large Filtrete",
                                    "Small Filtrete" = "Small Filtrete",
                                    "Winix" = c("Winix", "winix"))) %>% 
  mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
         kw_perc_expected = if_else(kw_hours == 1000, 999999, kw_perc_expected),
         kw_perc_expected = abs(kw_perc_expected)) %>% 
  replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) 
  filter(treatment_assigned == "Filter")

@@

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
pm_data <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
  arrange(area, home, pm_mean_sampling_period) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(area, home, pm_mean_sampling_period)
kids_kw_report_clean <- 
  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds"))

kw_perc_exp_summary <- kids_kw_report_clean %>% 
  left_join(pm_data, by = c("area", "home")) %>% 
  filter(!is.na(kw_perc_expected)) %>% 
  #filter(area == "WMT") %>% 
  group_by(filter_type, treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_expected),
            "SD KW % Exp" = sd(kw_perc_expected),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_expected),
            "Med KW % Exp" = median(kw_perc_expected),
            "Max KW % Exp" = max(kw_perc_expected),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
kw_perc_exp_summary

@@

# Try to extract sampling dates from KW hour data
```{r}
wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         sampling_date = ymd(samplingdate),
         pickup_date = ymd(pickupdate),
         area = "WMT") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

@@

@@

@@

kids_kw_filtered <- kids_kw_clean %>% 
  group_by(area, home_winter_id) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  mutate(max_kw_hours = if_else(kw_hours == max(kw_hours), 1, 0)) %>% 
  filter(max_kw_hours == 1) %>% 
  left_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
  mutate(date_diff = as.duration(interval(intervention_date, kw_date)),
         date_diff = as.numeric(date_diff)/86400) %>% 
  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
                              2.04, 0),
         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
                              4.224, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
                              0.888, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
                              3.48, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
                              0.0144, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
                              0.48, multiplier)) %>% 
  mutate(kw_expected = date_diff*multiplier,
         kw_perc_exp = kw_hours/kw_expected*100)

@@

>>>>

kids_kw_report_clean <- kids_kw_report %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
  #filter(winter_id == 1) %>% 
  arrange(area, home_winter_id) %>% 
  mutate(filter_type = fct_collapse(filter_type,
                                    "Honeywell" = "Honeywell",
                                    "Large Filtrete" = "Large Filtrete",
                                    "Small Filtrete" = "Small Filtrete",
                                    "Winix" = c("Winix", "winix"))) %>% 
  mutate(kw_perc_expected = if_else(kw_hours == -999, 999999, kw_perc_expected),
         kw_perc_expected = if_else(kw_hours == 1000, 999999, kw_perc_expected),
         kw_perc_expected = abs(kw_perc_expected)) %>% 
  replace_with_na(replace = list(kw_perc_expected = 999999)) %>% 
  mutate(kw_perc_expected = as.numeric(kw_perc_expected)) 
  filter(treatment_assigned == "Filter")

@@

# WMT filter unit compliance phone calls conducted by Carolyn
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

wmt_filter_checks <- read_xlsx(paste0(file_path, "Input/WMT/EdChecks-WithComments.xlsx"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, date, 
         filterrunning, 'filter details') %>% 
  mutate(home_winter_id = as.character(homewinterid),
         date = ymd(date),
         filter_level = as.factor(filterrunning),
         filter_comment = as.character(`filter details`),
         area = "WMT") %>% 
  select(home_winter_id, date, 
         filter_level, filter_comment, area) %>% 
  arrange(home_winter_id, date)

write_rds(wmt_filter_checks, paste0(file_path, "Output/wmt_filter_checks.rds"))
```

# Summarize KWhour and PM data; print table
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
pm_data <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds")) %>% 
  arrange(area, home, pm_mean_sampling_period) %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  dplyr::select(area, home, pm_mean_sampling_period, treatment_assigned, home_sqft_2level)
kids_kw_report_clean <- 
  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
  dplyr::select(-treatment_assigned)
wmt_filter_checks <- 
  read_rds(paste0(file_path, "Output/wmt_filter_checks.rds"))

kw_perc_exp_summary <- kids_kw_report_clean %>% 
  left_join(pm_data, by = c("area", "home")) %>% 
  right_join(wmt_filter_checks, by = c("area", "home_winter_id")) %>% 
  #filter(!is.na(kw_perc_expected)) %>% 
  #filter(treatment_assigned == "Filter") %>% 
  #distinct(area, home, .keep_all = TRUE) %>% 
  #filter(!is.na(filter_type)) %>% 
  #group_by(area, treatment_assigned, filter_type, home_sqft_2level) %>% 
  filter(area == "WMT") %>% 
  group_by(filter_type, treatment_assigned, filter_level) %>% 
  summarize("N KW % Exp" = n(),
            "Mean KW % Exp" = mean(kw_perc_expected, na.rm = TRUE),
            #"SD KW % Exp" = sd(kw_perc_expected, na.rm = TRUE),
            #"Min KW % Exp" = min(kw_perc_expected, na.rm = TRUE),
            "Med KW % Exp" = median(kw_perc_expected, na.rm = TRUE),
            #"Max KW % Exp" = max(kw_perc_expected),
            "Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "Med PM" = median(pm_mean_sampling_period, na.rm = TRUE))
kw_perc_exp_summary

# wmt_filte_checks data seems to validate the KW % expected as compliance data,
# at least for WMT. KW % Exp is lower in homes that reported filter as too low or off.

@@

# Extract sampling dates and use KWhour data from PM sampling period
```{r}
wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         sampling_date = ymd(samplingdate),
         pickup_date = ymd(pickupdate),
         area = "WMT") %>% 
  select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

@@

# Use KWhour data only if KWhours increased in subsequent dates
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

@@

filter_units <- 
  read_rds(paste0(file_path, "Output/kids_kw_report_clean.rds")) %>% 
  dplyr::select(home_winter_id, area, filter_type)

wmt_sampling_dates <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  dplyr::select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         sampling_date = ymd(samplingdate),
         pickup_date = ymd(pickupdate),
         area = "WMT") %>% 
  dplyr::select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

nn_sampling_dates <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  dplyr::select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "NN") %>% 
  dplyr::select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

ak_sampling_dates <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  dplyr::select(homewinterid, interventiondate, 
         samplingdate, pickupdate) %>% 
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         sampling_date = mdy(samplingdate),
         pickup_date = mdy(pickupdate),
         area = "AK") %>% 
  dplyr::select(home_winter_id, intervention_date, 
         sampling_date, pickup_date, area)

@@

kids_filter_compliance <- kids_kw_clean %>% 
  group_by(area, home_winter_id) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  filter(lag_diff >= 0 | is.na(lag_diff)) %>% 
  mutate(lag_diff = kw_hours - lag(kw_hours)) %>% 
  mutate(max_kw_hours = if_else(kw_hours == max(kw_hours), 1, 0)) %>% 
  filter(max_kw_hours == 1) %>% 
  left_join(kids_sampling_dates, by = c("area", "home_winter_id")) %>% 
  mutate(date_diff = as.duration(interval(intervention_date, kw_date)),
         date_diff = as.numeric(date_diff)/86400) %>% 
  left_join(filter_units, by = c("area", "home_winter_id")) %>% 
  mutate(multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Placebo",
                              2.04, 0),
         multiplier = if_else(filter_type == "Honeywell" & treatment_assigned == "Filter",
                              4.224, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Placebo",
                              0.888, multiplier),
         multiplier = if_else(filter_type == "Small Filtrete" & treatment_assigned == "Filter",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Placebo",
                              1.896, multiplier),
         multiplier = if_else(filter_type == "Large Filtrete" & treatment_assigned == "Filter",
                              3.48, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Placebo",
                              0.0144, multiplier),
         multiplier = if_else(filter_type == "Winix" & treatment_assigned == "Filter",
                              0.48, multiplier)) %>% 
  mutate(kw_expected = date_diff*multiplier,
         kw_perc_exp = kw_hours/kw_expected*100) %>% 
  dplyr::select(area, home_winter_id, winter_id, home, treatment_assigned, 
                kw_date, kw_hours, date_diff, filter_type, multiplier,
                kw_expected, kw_perc_exp)

compliance_summary1 <- kids_filter_compliance %>% 
  group_by(home) %>% 
  mutate(kw_perc_exp_total = mean(kw_perc_exp, na.rm = TRUE)) %>%  
  filter(treatment_assigned == "Filter") %>% 
  filter(!is.na(kw_perc_exp_total)) %>% 
  group_by(treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_exp_total),
            "SD KW % Exp" = sd(kw_perc_exp_total),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_exp_total),
            "25Q KW % Exp" = quantile(kw_perc_exp_total, 0.25),
            "Med KW % Exp" = median(kw_perc_exp_total),
            "75Q KW % Exp" = quantile(kw_perc_exp_total, 0.75),
            "Max KW % Exp" = max(kw_perc_exp_total))
compliance_summary1

kids_filter_compliance_total <- kids_filter_compliance %>% 
  group_by(home) %>% 
  mutate(kw_perc_exp_total = mean(kw_perc_exp), 
         treatment_assigned = as.character(treatment_assigned),
         perc_exp_quart = cut(kw_perc_exp_total,
                                  breaks = c(0, 30, 64, 100, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment_assigned == "Filter",
                                     perc_exp_quart, 
                                     treatment_assigned)) %>% 
  distinct(area, home, .keep_all = TRUE)

compliance_summary2 <- kids_filter_compliance %>% 
  group_by(area, home_winter_id) %>% 
  mutate(kw_perc_exp_winter = mean(kw_perc_exp, na.rm = TRUE)) %>%  
  filter(treatment_assigned == "Filter") %>% 
  filter(!is.na(kw_perc_exp_winter)) %>% 
  group_by(winter_id, treatment_assigned) %>% 
  summarize("Mean KW % Exp" = mean(kw_perc_exp_winter),
            "SD KW % Exp" = sd(kw_perc_exp_winter),
            "N KW % Exp" = n(),
            "Min KW % Exp" = min(kw_perc_exp_winter),
            "25Q KW % Exp" = quantile(kw_perc_exp_winter, 0.25),
            "Med KW % Exp" = median(kw_perc_exp_winter),
            "75Q KW % Exp" = quantile(kw_perc_exp_winter, 0.75),
            "Max KW % Exp" = max(kw_perc_exp_winter))
compliance_summary2

kids_filter_compliance_winter1 <- kids_filter_compliance %>% 
  filter(winter_id == 1) %>% 
  mutate(treatment_assigned = as.character(treatment_assigned),
         perc_exp_quart = cut(kw_perc_exp,
                                  breaks = c(0, 32, 60, 116, 1000),
                                  labels = c("0-25 compliance", "25-50 compliance",
                                             "50-75 compliance", "75-100 compliance")),
         perc_exp_quart = as.character(perc_exp_quart),
         filter_compliance = if_else(treatment_assigned == "Filter",
                                     perc_exp_quart, 
                                     treatment_assigned))

write_rds(kids_filter_compliance_total, paste0(file_path, "Output/kids_filter_compliance_total.rds"))
write_rds(kids_filter_compliance_winter1, paste0(file_path, "Output/kids_filter_compliance_winter1.rds"))

@@

# Changing the placebo multipliers above to the same as the filter multiplier
# changes the picture quite a bit. It seems likely that participants ran the 
# placebo filters on high more often than expected, resulting in high % Exp KWhours
